---
title: "importing_data_e1"
author: "Teodor Chakarov"
date: '2022-03-24'
output: 
  pdf_document:
    toc: TRUE
---
####################
# Tutorium in R 
####################
### Importing data with R
#### Exercise Introduction to Importing Data in R - Number 1
#### By: Teodor Chakarov 12141198

import csv
```{r}
pools <- read.csv("swimming_pools.csv")
str(pools)
```

stringsAsFactors convert strings columns in readable values
```{r}
pools = read.csv("swimming_pools.csv", stringsAsFactors=FALSE)
```


Import txt file, setting separator by TABS, removing the header and give us an overveiw of the data
```{r}
hotdogs = read.delim("hotdogs.txt", sep="\t", header = FALSE)
summary(hotdogs)
```

Show the path of the file and read the file
```{r}
path <- file.path("hotdogs.txt")
hotdogs <- read.table(path, 
                      sep = "", 
                      col.names = c("type", "calories", "sodium"))


```


Finish the read.delim() call, setting up the names of the columns
```{r}
hotdogs <- read.delim("hotdogs.txt", header = FALSE, col.names = c("type", "calories", "sodium"))
lily <- hotdogs[which.min(hotdogs$calories), ]
print(lily)
```
Select the observation with the most sodium: tom

```{r}
tom <- hotdogs[which.max(hotdogs$sodium), ]
print(tom)
```

```{r}
library(readr)

hotdogs2 <- read.delim("hotdogs.txt", header = FALSE, 
                       col.names = c("type", "calories", "sodium"),
                       colClasses = c("factor", "NULL", "numeric"))

properties <- c("area", "temp", "size", "storage", "method",
                "texture", "flavor", "moistness")

potatoes <- read_tsv("potatoes.txt", col_names=properties)

head(potatoes)

```


Skipping the first 2 observations

```{r}
potatoes_fragment <- read_tsv("potatoes.txt", skip = 2, n_max = 3, col_names = properties)
head(potatoes_fragment)
```

Set collector to for setting types of the columns
```{r}
fac <- col_factor(levels = c("Beef", "Meat", "Poultry"))
int <- col_integer()
hotdogs_factor <- read_tsv("hotdogs.txt",
                           col_names = c("type", "calories", "sodium"),
                           col_types = list(fac,int,int))

# Before
summary(hotdogs) #  type: Class :character  Mode  :character

# After
summary(hotdogs_factor)# type: Beef: 20, Meat: 17, Poultry: 17
```

## Visulization
```{r}

library(data.table) 
potatoes <- fread("potatoes.csv") # easy and fast to use for importing data
potatoes <- fread("potatoes.csv", select=c("texture", "moistness"))

# Plot texture (x) and moistness (y) of potatoes
plot(x=potatoes$texture, y=potatoes$moistness)
```
```
selecting only the columns we need

fread("path/to/file.txt", drop = 2:4)
fread("path/to/file.txt", select = c(1, 5))
fread("path/to/file.txt", drop = c("b", "c", "d"))
fread("path/to/file.txt", select = c("a", "e"))
```
The class of the result of fread() is both data.table and data.frame. read_csv() creates an object with three classes: tbl_df, tbl and data.frame

## Working with Excel
This package is on heavy development -> gdata is more reliable
```{r}
library(readxl)
excel_sheets("urbanpop.xlsx") # names of the sheets in the excel file
```
### Read the sheets, one by one
```{r}
pop_1 <- read_excel("urbanpop.xlsx", sheet = 1)
pop_2 <- read_excel("urbanpop.xlsx", sheet = 2)
pop_3 <- read_excel("urbanpop.xlsx", sheet = 3)
```

### Put pop_1, pop_2 and pop_3 in a list: pop_list
```{r}
pop_list <- list(pop_1, pop_2, pop_3)
str(pop_list)
```

Read all Excel sheets with lapply()
```{r}
pop_list <- lapply(excel_sheets("urbanpop.xlsx"), read_excel, path = "urbanpop.xlsx")

str(pop_list)
```

Setting column names using loop
```{r}
cols <- c("country", paste0("year_", 1960:1966))
pop_b <- read_excel("urbanpop_nonames.xlsx", col_names = cols)
```

Import the second sheet of urbanpop.xlsx, skipping the first 21 rows: urbanpop_sel
```{r}
urbanpop_sel <- read_excel("urbanpop.xlsx", skip=21, col_names=FALSE, sheet = 2)
head(urbanpop_sel, n=1)
```


```{r}
library(gdata)
perl<- "C:\\Users\\tedoc\\OneDrive\\Dokumente\\R\\win-library\\4.1\\rtools42\\usr\\bin\\perl5.32.1.exe"
urban_pop <- read.xls("urbanpop.xls", sheet = "1967-1974", perl = perl)
head(urban_pop, n = 11)
```

```{r}
columns <- c("country", paste0("year_", 1967:1974))


urban_pop <- read.xls("urbanpop.xls", sheet = 2,
                      skip = 50, header = FALSE, stringsAsFactors = FALSE,
                      col.names = columns, perl = perl)

head(urban_pop, n =10)

```


Import 3 different sheets
```{r message=FALSE, warning=FALSE}

path <- "urbanpop.xls"
urban_sheet1 <- read.xls(path, sheet = 1, stringsAsFactors = FALSE, perl = perl)
urban_sheet2 <- read.xls(path, sheet = 2, stringsAsFactors = FALSE, perl = perl)
urban_sheet3 <- read.xls(path, sheet = 3, stringsAsFactors = FALSE, perl = perl)
```

+ Combining the tables but dropping the first colums of 2 and 3

```{r}
urban <- cbind(urban_sheet1, urban_sheet2[-1], urban_sheet3[-1])

```

+ Remove all rows with NANs from urban
```{r}
urban_clean <- na.omit(urban)
```


## Build a "bridge" between your Excel file and your R session.
```{r}
library("XLConnect")

my_book <- loadWorkbook("urbanpop.xlsx")
class(my_book)
my_book <- loadWorkbook("urbanpop.xlsx")
```

List the sheets in my_book

```{r}
getSheets(my_book)

# Import the second sheet in my_book
readWorksheet(my_book, sheet = 2)
```


## Build connection to urbanpop.xlsx
```{r}
my_book <- loadWorkbook("urbanpop.xlsx")
```
Import columns 3, 4, and 5 from second sheet in my_book: urbanpop_sel

```{r}
urbanpop_sel <- readWorksheet(my_book, sheet = 2, startCol = 3, endCol = 5)
```

Import first column from second sheet in my_book: countries
```{r}
countries <- readWorksheet(my_book, sheet = 2, startCol = 1, endCol = 1)
```

cbind() urbanpop_sel and countries together: selection

```{r}
selection <- cbind(countries, urbanpop_sel)

# create new sheets
createSheet(my_book, name="data_summary")
```

```{r}
createSheet(my_book, "data_summary") # Add a worksheet to my_book, named "data_summary"
```

## Create data frame: summ
```{r}
sheets <- getSheets(my_book)[1:3]
dims <- sapply(sheets, function(x) dim(readWorksheet(my_book, sheet = x)), USE.NAMES = FALSE)
summ <- data.frame(sheets = sheets,
                   nrows = dims[1, ],
                   ncols = dims[2, ])
writeWorksheet(my_book, summ, "data_summary") # Add data in summ to "data_summary" sheet

saveWorkbook(my_book, "summary.xlsx") # Save workbook as summary.xlsx

renameSheet(my_book, "data_summary", "summary") # Rename "data_summary" sheet to "summary"

getSheets(my_book) # Print out sheets of my_book
```
```{r}
saveWorkbook(my_book, "renamed.xlsx")
removeSheet(my_book, 4)

saveWorkbook(my_book, "clean.xlsx") # Save workbook to "renamed.xlsx"

```
